---
title: "DS 6372"
author: "Oluwadamilola Owolabi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Calling Libraries
```{r}

#library(jtools)
library(sjPlot)
#library(ResourceSelection)
library(dplyr)
library(ggplot2)
```

Getting the data
```{r}

# Getting the csv file
#setwd("C:/Users/owola/Documents/MY_COURSES/SPRING_2024/DS_6372/Project/Project 2/bank+marketing/bank-additional/bank-additional")
Bank <-read.csv("C:/Users/owola/Documents/MY_COURSES/SPRING_2024/DS_6372/Project/Project 2/bank+marketing/bank-additional/bank-additional/Bank-additional-full.csv", header= TRUE, sep = ';')
head(Bank)
names(Bank)
```

Splitting the dataset into training and test Data
```{r}

# Split into training and test
set.seed(1234)
perc_train <- .8 # Should we make this higher?
train_indices <- sample(nrow(Bank), floor(perc_train * nrow(Bank)))
BankData <- Bank[train_indices, ] #the training dataset
BankTest <- Bank[-train_indices, ] #the validation dataset.
yes_data <- BankTest[BankData$y == 'yes',] # 4250
no_data <- BankData[BankTest$y == 'no',] # 31918

BankData$y.num<-ifelse(BankData$y=="yes",1,0)
BankData
```




#### Looking at logstic regression P-values on just the numerical variables
 
```{r}

log.model.numerical <-glm(y.num ~ age + duration + campaign + pdays + previous + emp.var.rate + cons.price.idx + cons.conf.idx + euribor3m + nr.employed ,data = BankData,family="binomial")
coef(summary(log.model.numerical))
```
variables with p-values < 0.05, it means the variable is statistically significant
Variables with statistical significance :- pdays, previous, emp.var.rate, cons.price.idx, cons.conf.idx, nr.employed.

Creating a bar graph to show the intensity of each p-value

```{r}

data <- c(9.845997e-110, 7.352136e-12, 3.264951e-21, 4.943288e-10, 2.273849e-07, 1.082911e-06)  # Numeric data for bar heights
log_data <- log(data) * -1
labels <- c('pdays', 'previous', 'emp.var.rate', 'cons.price.idx', 'cons.conf.idx', 'nr.employed')  # Labels for the categories

# Create a bar plot
barplot(log_data, names.arg = labels, 
        xlab = "Categories", ylab = "Frequency/Counts",
        main = "Bar Plot Example",
        col = "skyblue",  # Bar color
        border = "black",  # Border color
        ylim = c(0, max(log_data) * 1))  # Adjust y-axis limits for better visualization
```


#### Logistic regressions for categorical variables

```{r}

log.model.categorical <-glm(y.num ~ job + marital + education + default + housing + loan + contact + month + day_of_week + poutcome ,data = BankData,family="binomial")
coef(summary(log.model.categorical))
```
## variables with p-values < 0.05, it means the variable is statistically significant
## Variables with statistical significance :- jobblue-collar, jobhousemaid, jobretired, jobservices, jobstudent, jobtechnician, maritalsingle, jobunemployed, maritalsingle, educationunknown, defaultunknown, contacttelephone, monthaug, monthdec, monthjul, monthmar, monthmay, monthnov, monthoct, monthsep, day_of_weekmon, day_of_weekwed, poutcomesuccess 

Creating a bar graph to show the intensity of each p-value

```{r}

#Had to scrap this code. Too much hard coding

# data <- c(1.551613e-33, 5.995821e-01, 1.592811e-14, 4.456974e-03, 1.195754e-10, 3.376630e-02, 5.436889e-03, 4.942619e-02, 4.420344e-22, 7.001195e-78, 1.497726e-34, 1.092251e-06, 5.133784e-30, 4.295017e-27, 3.491645e-34, 1.036440e-31, 5.489694e-15, 1.572625e-03,  1.163557e-07, 2.343290e-02, 3.047043e-185)  # Numeric data for bar heights
# log_data <- log(data) * -1
# labels <- c('jobblue-collar', 'jobhousemaid', 'jobretired', 'jobservices', 'jobstudent', 'jobtechnician', 'maritalsingle', 'jobunemployed', 'maritalsingle', 'educationunknown', 'defaultunknown', 'contacttelephone', 'monthaug', 'monthdec', 'monthjul', 'monthmar', 'monthmay', 'monthnov', 'monthoct', 'monthsep', 'day_of_weekmon', 'day_of_weekwed', 'poutcomesuccess')  # Labels for the categories
# 
# # Create a bar plot
# barplot(log_data, names.arg = labels, 
#         xlab = "Categories", ylab = "Frequency/Counts",
#         main = "Bar Plot Example",
#         col = "skyblue",  # Bar color
#         border = "black",  # Border color
#         ylim = c(0, max(log_data) * 1))  # Adjust y-axis limits for better visualization
```

```{r}
names(BankData)

log.model <-glm(y.num ~ job + marital + education + default + housing + loan + contact + month + day_of_week + poutcome + age + duration + campaign + pdays + previous + emp.var.rate + cons.price.idx + cons.conf.idx + euribor3m + nr.employed ,data = BankData,family="binomial")
coef(summary(log.model))
```

```{r}
# Extract variable names
variable_names <- rownames(summary(log.model)$coefficients)

# getting the  p-values from the model3
p_values <- summary(log.model)$coefficients[, 4]  # Assuming p-values are in the 5th column of the summary table
p_values <- data.frame(p_values)$p_values

df <- data.frame(variable_names, p_values) #combining the pvalues and variable names into a dataframe

df <- df[!df$p_value == 0 , ] #removing varaiables with pvalue = 0
df$p_values <- log(df$p_values)  * -1

barplot(df$p_values, 
        main = "P-values of Regression Coefficients less than the significance level 0.05", 
        xlab = "Variables", 
        ylab = "P-value",
        names.arg = df$variable_names,
        las = 2,  # Rotate x-axis labels vertically for better readability
        col = "steelblue",  # Set color of bars
        ylim = c(exp(0.05) * -1, max(df$p_values) * 1.2)  # Set ylim from the significance level to the minimum pvalue
        
)

library(ggplot2)
ggplot(df,aes(variable_names,p_values, fill = ifelse(p_values > (exp(0.05) * -1), "Positive", "Negative"))) +
  geom_bar(stat="identity", fill = "skyblue") + 
  #geom_text(aes(label = variable_names), vjust = -0.5) +  # Add text labels on top of bars
  scale_fill_manual(values = c("Positive" = "skyblue", "Negative" = "salmon")) +
  labs(x = "Variables", y = "P-values", title = "P-values of Regression Coefficients less than the significance level 0.05") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))  # Rotate x-axis labels for better readability

```

## Using forward/backward model selection

Forward selection
```{r}
library(leaps)

model.fwd = regsubsets(y.num ~ job + marital + education + default + housing + loan + contact + month + day_of_week + poutcome + age + duration + campaign + pdays + previous + emp.var.rate + cons.price.idx + cons.conf.idx + euribor3m + nr.employed ,data = BankData, nvmax=20, method="forward")

summary(model.fwd)$adjr2
summary(model.fwd)$rss
summary(model.fwd)$bic

#par(mar=c(.000001,.000001,.000001,.000001))
bics<-summary(model.fwd)$bic
plot(1:21,bics,type="l",ylab="BIC",xlab="# of predictors") #got 21 from the forward selection prediction
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=10)

adjr2<-summary(model.fwd)$adjr2
plot(1:21,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=10)

rss<-summary(model.fwd)$rss
plot(1:21,rss,type="l",ylab="train RSS",xlab="# of predictors")
index<-which(rss==min(rss))
points(index,rss[index],col="red",pch=10)

coef(model.fwd,21) #looking at coefficints

#Ans: out of 10 categorical variables and 10 munerical variables, the feature selection chose 14 subcategorical variables and 7 numerical variables.

# The variables are "jobblue-collar", "jobretired", "jobstudent", "monthaug", "monthdec", "monthjul", "monthjun", "monthmar", "monthmay", "monthnov", "day_of_weekmon", "day_of_weekthu", "poutcomesuccess", "age", "campaign", "previous", "cons.price.idx", "cons.conf.idx", "euribor3m", "nr.employed", "loanunknown".
```

backward selection
```{r}

model.bwd=regsubsets(y.num ~ job + marital + education + default + housing + loan + contact + month + day_of_week + poutcome + age + duration + campaign + pdays + previous + emp.var.rate + cons.price.idx + cons.conf.idx + euribor3m + nr.employed ,data = BankData, nvmax=20, method="backward", really.big=TRUE)

summary(model.bwd)$adjr2
summary(model.bwd)$rss
summary(model.bwd)$bic

#par(mar=c(.01,.01,.01,.01))
#par(mfrow=c(1,21))
bics<-summary(model.bwd)$bic
plot(1:21,bics,type="l",ylab="BIC",xlab="# of predictors") #got 7 from the forward selection prediction
index<-which(bics==min(bics))
points(index,bics[index],col="red",pch=30)

adjr2<-summary(model.bwd)$adjr2
plot(1:21,adjr2,type="l",ylab="Adjusted R-squared",xlab="# of predictors")
index<-which(adjr2==max(adjr2))
points(index,adjr2[index],col="red",pch=30)

rss<-summary(model.bwd)$rss
plot(1:21,rss,type="l",ylab="train RSS",xlab="# of predictors")
index<-which(rss==min(rss))
points(index,rss[index],col="red",pch=30)

coef(model.fwd,21) #looking at coefficints

#Ans: out of 10 categorical variables and 10 munerical variables, the feature selection chose 14 subcategorical variables and 7 numerical variables.

# The variables are "jobblue-collar", "jobretired", "jobstudent", "monthaug", "monthdec", "monthjul", "monthjun", "monthmar", "monthmay", "monthnov", "day_of_weekmon", "day_of_weekthu", "poutcomesuccess", "age", "campaign", "previous", "cons.price.idx", "cons.conf.idx", "euribor3m", "nr.employed", "loanunknown".

```

