---
title: "Final Project - "
author: "Aaron Abromowitz, Stephanie Duarte, Dammy "
date: "2024-04-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# EDA
The first thing that is done when you get a dataset is to peform an Exploratory Data Anslysis, to see what characteristics the data has.  The variable we are trying to predict is the y column whic represents if the client has subscribed to a term deposit.  The possible values are "yes" or "no".

## Create training and test set
Going forward, we will use the training set for all analysis and model building.  The test set will be used at the end to get metrics for the various models we create.
```{r}
# Pull in data
data<-read.csv("C:/Users/aabro/OneDrive/Desktop/SMU Program/Classes/Stats 2/Final Project/bank-additional-full.csv",stringsAsFactors = T, sep=";")

# Set levels to use for later
data$y <- relevel(data$y, ref="yes")
data$month <- factor(data$month, levels=c('mar','apr','may','jun','jul','aug','sep','oct','nov','dec'))
data$day_of_week <- factor(data$day_of_week, levels=c('mon','tue','wed','thu','fri'))

# Create the train and test split
train_perc <- .8
set.seed(1234)
train_indices <- sample(nrow(data), floor(train_perc * nrow(data)))
train_data <- data[train_indices, ]
nrow(train_data)
test_data <- data[-train_indices, ] 
nrow(test_data)
train_data$duration <- c()
```

## Look at summary statistics for numeric variables
There are several numeric variables where we can look at the min/max, quartiles, median, and mean.
```{r}
summary(train_data$age)
summary(train_data$campaign)
summary(train_data$pdays)
summary(train_data$previous)
summary(train_data$emp.var.rate)
summary(train_data$cons.price.idx)
summary(train_data$cons.conf.idx)
summary(train_data$euribor3m)
summary(train_data$nr.employed)
```

## Look at summary statistics for categorical variables
There are several categorical variables.  Summary statistics don't make as much sense for them, but you can look at the distribution of values in the different categories.
```{r}
summary(train_data$job)
summary(summary(train_data$job))
length(summary(train_data$job))

summary(train_data$marital)
summary(summary(train_data$marital))
length(summary(train_data$marital))

summary(train_data$education)
summary(summary(train_data$education))
length(summary(train_data$education))

summary(train_data$default)
summary(summary(train_data$default))
length(summary(train_data$default))

summary(train_data$housing)
summary(summary(train_data$housing))
length(summary(train_data$housing))

summary(train_data$loan)
summary(summary(train_data$loan))
length(summary(train_data$loan))

summary(train_data$contact)
summary(summary(train_data$contact))
length(summary(train_data$contact))

summary(train_data$month)
summary(summary(train_data$month))
length(summary(train_data$month))

summary(train_data$day_of_week)
summary(summary(train_data$day_of_week))
length(summary(train_data$day_of_week))

summary(train_data$poutcome)
summary(summary(train_data$poutcome))
length(summary(train_data$poutcome))
```

## Examine bank client data
The dataset includes age, job, marital, education, default, housing, and loan columns, which were identified as client data.

```{r}
library(tidyverse)

# Plot age
summary <- train_data %>%
  group_by(age,y) %>%
  summarize(count=n())
summary$perc <- 0
summary$perc[summary$y == 'no'] <- summary$count[summary$y == 'no'] / nrow(train_data[train_data$y == 'no',]) * 100
summary$perc[summary$y == 'yes'] <- summary$count[summary$y == 'yes'] / nrow(train_data[train_data$y == 'yes',]) * 100
summary %>% ggplot(aes(x=age,y=perc,fill=y)) + geom_bar(stat="identity") + facet_wrap(~y) + 
  ylab('Percentage') + xlab('Age') + ggtitle('Age Distribution Based on Y Value')

# Plot job
summary <- train_data %>%
  group_by(job,y) %>%
  summarize(count=n())
summary$perc <- 0
summary$perc[summary$y == 'no'] <- summary$count[summary$y == 'no'] / nrow(train_data[train_data$y == 'no',]) * 100
summary$perc[summary$y == 'yes'] <- summary$count[summary$y == 'yes'] / nrow(train_data[train_data$y == 'yes',]) * 100
summary %>% ggplot(aes(x=job,y=perc,fill=y)) + geom_bar(stat="identity") + facet_wrap(~y) + 
  ylab('Percentage') + xlab('Job') + ggtitle('Job Percentages Based on Y Value') + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot marital
summary <- train_data %>%
  group_by(marital,y) %>%
  summarize(count=n())
summary$perc <- 0
summary$perc[summary$y == 'no'] <- summary$count[summary$y == 'no'] / nrow(train_data[train_data$y == 'no',]) * 100
summary$perc[summary$y == 'yes'] <- summary$count[summary$y == 'yes'] / nrow(train_data[train_data$y == 'yes',]) * 100
summary %>% ggplot(aes(x=marital,y=perc,fill=y)) + geom_bar(stat="identity") + facet_wrap(~y) + 
  ylab('Percentage') + xlab('Marital') + ggtitle('Marital Percentages Based on Y Value') + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot default
summary <- train_data %>%
  group_by(default,y) %>%
  summarize(count=n())
summary$perc <- 0
summary$perc[summary$y == 'no'] <- summary$count[summary$y == 'no'] / nrow(train_data[train_data$y == 'no',]) * 100
summary$perc[summary$y == 'yes'] <- summary$count[summary$y == 'yes'] / nrow(train_data[train_data$y == 'yes',]) * 100
summary %>% ggplot(aes(x=default,y=perc,fill=y)) + geom_bar(stat="identity") + facet_wrap(~y) + 
  ylab('Percentage') + xlab('Default') + ggtitle('Default Percentages Based on Y Value') + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot housing
summary <- train_data %>%
  group_by(housing,y) %>%
  summarize(count=n())
summary$perc <- 0
summary$perc[summary$y == 'no'] <- summary$count[summary$y == 'no'] / nrow(train_data[train_data$y == 'no',]) * 100
summary$perc[summary$y == 'yes'] <- summary$count[summary$y == 'yes'] / nrow(train_data[train_data$y == 'yes',]) * 100
summary %>% ggplot(aes(x=housing,y=perc,fill=y)) + geom_bar(stat="identity") + facet_wrap(~y) + 
  ylab('Percentage') + xlab('Housing') + ggtitle('Housing Percentages Based on Y Value') + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
<br/>
None of these variables appear to show any strong indicator for yes or no.  It looks like higher ages tend to lean more towards yes, and certain jobs (Ex: admin) lean more towards yes.  When Y = no, there are more about double the unknown values, but still far under 50%.

## Examine data related with the last contact of the current campaign
The dataset includes contact communication type, month of contact, and day of week of contac, for the last contact of the current campaign to sell term deposits.
```{r}
# Plot contact
summary <- train_data %>%
  group_by(contact,y) %>%
  summarize(count=n())
summary$perc <- 0
summary$perc[summary$y == 'no'] <- summary$count[summary$y == 'no'] / nrow(train_data[train_data$y == 'no',]) * 100
summary$perc[summary$y == 'yes'] <- summary$count[summary$y == 'yes'] / nrow(train_data[train_data$y == 'yes',]) * 100
summary %>% ggplot(aes(x=contact,y=perc,fill=y)) + geom_bar(stat="identity") + facet_wrap(~y) + 
  ylab('Percentage') + xlab('Contact') + ggtitle('Contact Percentages Based on Y Value') + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot month
summary <- train_data %>%
  group_by(month,y) %>%
  summarize(count=n())
summary$perc <- 0
summary$perc[summary$y == 'no'] <- summary$count[summary$y == 'no'] / nrow(train_data[train_data$y == 'no',]) * 100
summary$perc[summary$y == 'yes'] <- summary$count[summary$y == 'yes'] / nrow(train_data[train_data$y == 'yes',]) * 100
summary %>% ggplot(aes(x=month,y=perc,fill=y)) + geom_bar(stat="identity") + facet_wrap(~y) + 
  ylab('Percentage') + xlab('Month') + ggtitle('Month Percentages Based on Y Value') + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot day of week
summary <- train_data %>%
  group_by(day_of_week,y) %>%
  summarize(count=n())
summary$perc <- 0
summary$perc[summary$y == 'no'] <- summary$count[summary$y == 'no'] / nrow(train_data[train_data$y == 'no',]) * 100
summary$perc[summary$y == 'yes'] <- summary$count[summary$y == 'yes'] / nrow(train_data[train_data$y == 'yes',]) * 100
summary %>% ggplot(aes(x=day_of_week,y=perc,fill=y)) + geom_bar(stat="identity") + facet_wrap(~y) + 
  ylab('Percentage') + xlab('Day of Week') + ggtitle('Day of Week Percentages Based on Y Value') + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
If y is Yes, there are ~20% more likely to be contacted on your cell phone than on landline.  There are certain months that also seem to have more term deposit sales in them.  Day of week looks to not have much change.

## Examine other data related with the current campaign or previous campaigns
```{r}
# Plot campaign
summary <- train_data %>%
  group_by(campaign,y) %>%
  summarize(count=n())
summary$perc <- 0
summary$perc[summary$y == 'no'] <- summary$count[summary$y == 'no'] / nrow(train_data[train_data$y == 'no',]) * 100
summary$perc[summary$y == 'yes'] <- summary$count[summary$y == 'yes'] / nrow(train_data[train_data$y == 'yes',]) * 100
summary %>% ggplot(aes(x=campaign,y=perc,fill=y)) + geom_bar(stat="identity") + facet_wrap(~y) + 
  ylab('Percentage') + xlab('Campaign') + ggtitle('Campaign Distribution Based on Y Value') + xlim(c(0,20))

# Plot poutcome
summary <- train_data %>%
  group_by(poutcome,y) %>%
  summarize(count=n())
summary$perc <- 0
summary$perc[summary$y == 'no'] <- summary$count[summary$y == 'no'] / nrow(train_data[train_data$y == 'no',]) * 100
summary$perc[summary$y == 'yes'] <- summary$count[summary$y == 'yes'] / nrow(train_data[train_data$y == 'yes',]) * 100
summary %>% ggplot(aes(x=poutcome,y=perc,fill=y)) + geom_bar(stat="identity") + facet_wrap(~y) + 
  ylab('Percentage') + xlab('Poutcome') + ggtitle('Poutcome Percentages Based on Y Value') + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Campaign distributions seem pretty similar between Yes and No.  For Poutcome, success is more common for the yes than for no.

