---
title: "Final Project - "
author: "Aaron Abromowitz, Stephanie Duarte, Dammy "
date: "2024-04-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# EDA
The first thing that is done when you get a dataset is to peform an Exploratory Data Anslysis, to see what characteristics the data has.  The variable we are trying to predict is the y column whic represents if the client has subscribed to a term deposit.  The possible values are "yes" or "no".

## Create training and test set
Going forward, we will use the training set for all analysis and model building.  The test set will be used at the end to get metrics for the various models we create.
```{r}
# Pull in data
data<-read.csv("C:/Users/aabro/OneDrive/Desktop/SMU Program/Classes/Stats 2/Final Project/bank-additional-full.csv",stringsAsFactors = T, sep=";")

# Set levels to use for later
data$y <- relevel(data$y, ref="yes")
data$month <- factor(data$month, levels=c('mar','apr','may','jun','jul','aug','sep','oct','nov','dec'))

# Create the train and test split
train_perc <- .8
set.seed(1234)
train_indices <- sample(nrow(data), floor(train_perc * nrow(data)))
train_data <- data[train_indices, ]
nrow(train_data)
test_data <- data[-train_indices, ] 
nrow(test_data)
train_data$duration <- c()
```

## Look at summary statistics for numeric variables
There are several numeric variables where we can look at the min/max, quartiles, median, and mean.

## Examine bank client data
The dataset includes age, job, marital, education, default, housing, and loan columns, which were identified as client data.

```{r}
library(tidyverse)

# Plot age
summary <- train_data %>%
  group_by(age,y) %>%
  summarize(count=n())
summary$perc <- 0
summary$perc[summary$y == 'no'] <- summary$count[summary$y == 'no'] / nrow(train_data[train_data$y == 'no',]) * 100
summary$perc[summary$y == 'yes'] <- summary$count[summary$y == 'yes'] / nrow(train_data[train_data$y == 'yes',]) * 100
summary %>% ggplot(aes(x=age,y=perc,fill=y)) + geom_bar(stat="identity") + facet_wrap(~y) + 
  ylab('Percentage') + xlab('Age') + ggtitle('Age Distribution Based on Y Value')

# Plot job
summary <- train_data %>%
  group_by(job,y) %>%
  summarize(count=n())
summary$perc <- 0
summary$perc[summary$y == 'no'] <- summary$count[summary$y == 'no'] / nrow(train_data[train_data$y == 'no',]) * 100
summary$perc[summary$y == 'yes'] <- summary$count[summary$y == 'yes'] / nrow(train_data[train_data$y == 'yes',]) * 100
summary %>% ggplot(aes(x=job,y=perc,fill=y)) + geom_bar(stat="identity") + facet_wrap(~y) + 
  ylab('Percentage') + xlab('Job') + ggtitle('Job Percentages Based on Y Value') + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot marital
summary <- train_data %>%
  group_by(marital,y) %>%
  summarize(count=n())
summary$perc <- 0
summary$perc[summary$y == 'no'] <- summary$count[summary$y == 'no'] / nrow(train_data[train_data$y == 'no',]) * 100
summary$perc[summary$y == 'yes'] <- summary$count[summary$y == 'yes'] / nrow(train_data[train_data$y == 'yes',]) * 100
summary %>% ggplot(aes(x=marital,y=perc,fill=y)) + geom_bar(stat="identity") + facet_wrap(~y) + 
  ylab('Percentage') + xlab('Marital') + ggtitle('Marital Percentages Based on Y Value') + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
<br/>
None of these variables appear to show any strong indicator for yes or no.  It looks like higher ages tend to lean more towards yes, and certain jobs (Ex: admin) lean more towards yes.  But none of these variables look particularly strong.